A Partitioning Algorithm for Maximum Common Subgraph Problems\thanks{This work was supported by a pile of money from a funding agency.}

### 2017-02-15 paper/paper.tex
Some things that changed

% TODO: use operatorname for these
\DeclareMathOperator{\V}{V}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\N}{N}
\DeclareMathOperator{\vtxlabel}{label}

@@

\section{Preliminaries}

We initially assume that graphs are unlabelled, undirected and without loops
(Section~\ref{sec:extensions} describes how these restrictions may be relaxed).
The vertices and edge sets of a graph $G$ are denoted $\V(G)$ and $\E(G)$.  The
set of vertices adjacent to vertex $v$ in graph $G$ is called the
\emph{neighbourhood} of $v$, and denoted $\N(v, G)$.  The graph $\overline{G}$
is the complement of $G$, obtained copying the vertices of $G$ and adding an
edge between each pair of distinct vertices if and only if $G$ does not
have an edge between the corresponding pair.

@@

\section{Prior Work}

The most successful existing approaches to solving MCS are reformulation to the
maximum clique problem on one hand \cite{DBLP:conf/cp/McCreeshNPS16}, and
constraint programming on the other \cite{DBLP:conf/cp/NdiayeS11}. In the
clique encoding, an association graph is created with one node for each pair in
the set $\{(u,v) \mid u \in \V(G), v \in \V(H) \}$ and an edge for each pair of
mappings that may be chosen together. A maximum clique in this graph
corresponds directly to a maximum common subgraph of $G$ and $H$.

In the constraint programming approach  \cite{DBLP:conf/cp/NdiayeS11}, a variable corresponds to
each vertex in $G$, and each variable initially has one value in its domain for
each vertex in $H$ to which it may be mapped, plus an additional value $\bot$
signifying that the vertex remains unmapped. A backtracking search is performed
in which values are assigned to variables, and propagation rules are used to
remove infeasible values from domains.
\textit{A matching algorithm propagates
a soft global allDifferent constraint which maximizes the number variables that are assigned to values different from $\bot$, while ensuring they are all different when they are not assigned to $\bot$.
Different constraint propagation techniques are compared in \cite{DBLP:conf/cp/NdiayeS11}. The combination ``MAC+Bound'' (resp. ``FC+Bound'') obtains the best results on labelled (resp. non labelled) graphs and outperforms the approach of \cite{McGreg82} and the CP model of \cite{DBLP:conf/mco/VismaraV08}. The combination ``MAC+Bound'' maintains arc consistency~\cite{sabi94} of edge constraints, whereas the conbination ``FC+Bound'' simply performs forward-checking on these constraints. In both combinations, the ``Bound'' filtering checks whether it is possible to assign distinct values to enough variables to surpass the best cost found so far (it is a weaker version of GAC(\emph{softAllDiff}) \cite{peti01} which computes the maximum number of variables that can be assigned distinct values).}

We now present \McSplit, a branch and bound algorithm 
which uses a new vertex-labelling scheme to avoid explicitly maintaining the
set of feasible mappings for each vertex in $\V(G)$. 
\iffalse
As we will explore in
Section~\ref{sec:comparison}, the algorithm may be viewed as a version of the
constraint programming forward-checking (CP-FC) algorithm using a compact data
structure that exploits the nature of the MCS problem.  It is sufficiently
different, however, that it is helpful to describe the algorithm without
reference to previous approaches.

@@

\begin{algorithm}[ht]
\DontPrintSemicolon
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
\nl \Begin{
%\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$save(matching)$} \label{StoreIncumbent}
%\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
\nl $bf \gets 0$ \label{SetBestFutureToZero} \;
\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}{$\AlgVar{bf} \gets \AlgVar{bf} + \min(|P|,|T|)$}
\nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
\bigskip
\nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
\nl $v \gets \FuncSty{SelectVertex}(P)$ \label{SelectVertex} \;
\nl \For {$w \in T$ \label{WLoop}} {
\nl    $M \gets M \cup (v,w)$ \label{GrowM} \;
\nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
\nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
\nl        \bf{Let} $P'' \gets P' \cap N(v,G) \setminus \{v\}$ \label{NewPWithEdge} \;
\nl        \bf{Let} $T'' \gets T' \cap N(w,H) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
\nl        \bf{Let} $P'' \gets P' \cap N(v,\overline{G}) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
\nl        \bf{Let} $T'' \gets T' \cap N(w,\overline{H}) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
       }
\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
\nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
  }
\nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\}$\;
\nl \lIf {$P' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle P',T \rangle \}$}
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
}
\;
\nl $\FuncSty{McSplit}(G,H)$ \label{McSplitFun} \;
\nl \Begin{
\nl $\AlgVar{incumbent} \gets \emptyset$ \;
\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
\nl return $|\AlgVar{incubent}|$ \;
}
\caption{\McSplit\ algorithm for Maximum Common Subgraph}
\label{McCliqueAlg}
\end{algorithm}

@@

\section{Comparison with Existing Algorithms}\label{sec:comparison} The
\McSplit\ algorithm may be considered as a version of the constraint
programming, forward-checking (CP-FC) algorithm of
\citep{DBLP:conf/cp/NdiayeS11}, using an efficient, compressed representation
of domains.  (Recall that in the CP-FC algorithm, each vertex in $v \in \V(G)$
is represented by a variable, whose domain corresponds to the set of vertices
in $\V(H)$ to which $v$ may currently be mapped.)

@@

\begin{algorithm}
\DontPrintSemicolon
\nl    \For {$u \in \N(b,G)$}{
\nl        remove $v$ from $D'_u$ \;
\nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
       }
\nl    \For {$u \in \N(b,\overline{G})$}{
\nl        remove $v$ from $D'_u$ \;
\nl        remove the neighbourhood of $t$ from $D'_u$ \;
       }
\caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm}
\label{cpAlg}
\end{algorithm}

>>>>

% TODO: use operatorname for these
\DeclareMathOperator{\V}{V}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\N}{N}
\DeclareMathOperator{\invN}{\overline{N}}
\DeclareMathOperator{\vtxlabel}{label}

@@

We initially assume that graphs are unlabelled, undirected and without loops
(Section~\ref{sec:extensions} describes how these restrictions may be relaxed).
The vertices and edge sets of a graph $G$ are denoted $\V(G)$ and $\E(G)$.  The
set of vertices adjacent to vertex $v$ in graph $G$ is called the
\emph{neighbourhood} of $v$, denoted $\N(G, v)$. We denote by $\invN(G, v)$ the
\emph{inverse neighbourhood} of $v$, being the set of the vertices not adjacent
to $v$ (excluding $v$ itself).

@@

With these definitions established, we now present \McSplit, a branch and bound
algorithm which uses a new vertex-labelling scheme to avoid explicitly
maintaining the set of feasible mappings for each vertex in $\V(G)$.
\iffalse
As we will explore in
Section~\ref{sec:comparison}, the algorithm may be viewed as a version of the
constraint programming forward-checking (CP-FC) algorithm using a compact data
structure that exploits the nature of the MCS problem.  It is sufficiently
different, however, that it is helpful to describe the algorithm without
reference to previous approaches.

@@

\begin{algorithm}[ht]
\DontPrintSemicolon
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
\nl \Begin{
%\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$save(matching)$} \label{StoreIncumbent}
%\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
\nl $bf \gets 0$ \label{SetBestFutureToZero} \;
\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}{$\AlgVar{bf} \gets \AlgVar{bf} + \min(|P|,|T|)$}
\nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
\bigskip
\nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
\nl $v \gets \FuncSty{SelectVertex}(P)$ \label{SelectVertex} \;
\nl \For {$w \in T$ \label{WLoop}} {
\nl    $M \gets M \cup (v,w)$ \label{GrowM} \;
\nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
\nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
\nl        \bf{Let} $P'' \gets P' \cap \N(G, G) \setminus \{v\}$ \label{NewPWithEdge} \;
\nl        \bf{Let} $T'' \gets T' \cap \N(H, w) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
\nl        \bf{Let} $P'' \gets P' \cap \invN(G, v) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
\nl        \bf{Let} $T'' \gets T' \cap \invN(H, w) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
       }
\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
\nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
  }
\nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\}$\;
\nl \lIf {$P' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle P',T \rangle \}$}
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
}
\;
\nl $\FuncSty{McSplit}(G,H)$ \label{McSplitFun} \;
\nl \Begin{
\nl $\AlgVar{incumbent} \gets \emptyset$ \;
\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
\nl return $|\AlgVar{incubent}|$ \;
}
\caption{\McSplit\ algorithm for Maximum Common Subgraph}
\label{McCliqueAlg}
\end{algorithm}

@@

\section{Comparison with Existing Algorithms}\label{sec:comparison} The
\McSplit\ algorithm may be considered as a version of the constraint
programming, forward-checking (CP-FC) algorithm of
\citep{DBLP:conf/cp/NdiayeS11}, using an efficient, compressed representation
of domains.  In the CP-FC algorithm, each vertex in $v \in \V(G)$
is represented by a variable, whose domain corresponds to the set of vertices
in $\V(H)$ to which $v$ may currently be mapped.

@@

\begin{algorithm}
\DontPrintSemicolon
\nl    \For {$u \in \N(G, b)$}{
\nl        remove $v$ from $D'_u$ \;
\nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
       }
\nl    \For {$u \in \invN(G, b)$}{
\nl        remove $v$ from $D'_u$ \;
\nl        remove the neighbourhood of $t$ from $D'_u$ \;
       }
\caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm}
\label{cpAlg}
\end{algorithm}
